### 2008-12-16 talk/ecoop2009/clibackend.tex

A flexswitch is a special kind of switch which can be dynamically
extended with new cases; intuitively, its behavior can be described
well in terms of flow graphs. Indeed, a flexswitch can be considered 
as a special flow graph block where links to newly created blocks are
dynamically added whenever new cases are needed. \dacom{in theory links could also be removed, but I
  do not know whether this possibility has been considered in Rainbow}

@@

Implementing flexswitches for backends generating assembly code is
quite straightforward: basically, a new jump has to be inserted in the
existing code to point to the newly generated code fragment.

@@

Since in .NET methods are the basic units of compilation, a possible
solution consists in creating a new method 
any time a new case has to be added to a flexswitch.
\dacom{comment for Antonio: I am not sure this is the best solution. This cannot work for Java where classes are the basic
  units}

\dacom{I still have to polish what comes next}

\newcommand{\commentout}[1]{}
\commentout{
Because of all these constraints we cannot simply map each graph to its own method, since we saw that our graphs can grow after they have already been executed few times.

Hence, we need to distinguish between the two concepts:

        * a graph is the logical unit of code as seen by the JIT compiler: concretely, the CLI JIT backend renders it as one or more methods;
        * a method is a collection of basic blocks; each method has the so called parent graph, i.e. the graph its blocks logically belongs to.

The first method of a graph is called main method (which has nothing to do with the Main static methods found in .exe files); other methods are called children methods.

When we want to add a new case to the flexswitch, we create a method containing all the new code; then we wrap the method inside a delegate (the .NET equivalent of a function pointer) and pass it to the flexswitch, so that it can later invoke it.
The hard bit: non-local links

@@

Moreover, the dispatch blocks are emitted only if needed, i.e. if the parent graph contains at least one flexswitch; graphs without flexswitches are rendered in the obvious way, by making one method per graph.
The slow bit: passing arguments

@@

>>>>

A flexswitch is a special kind of switch which can be dynamically
extended with new cases; intuitively, its behavior can be described
well in terms of flow graphs. Indeed, a flexswitch can be considered 
as a special flow graph block where links to newly created blocks are
dynamically added whenever new cases are needed.

@@

Implementing flexswitches for backends generating machine code is
not too complex: basically, a new jump has to be inserted in the
existing code to point to the newly generated code fragment.

@@

Since in .NET methods are the basic units of compilation, a possible
solution consists in creating a new method 
any time a new case has to be added to a flexswitch.
\dacom{comment for Antonio: I am not sure this is the best solution. This cannot work for Java where classes are the basic
  units. Closures will be available only with Java Dolphin and I do
  not know how much efficient will be}
In this way, whereas flow graphs without flexswitches are translated
to a single method, the translation of flow graphs which can dynamically grow because of
flexswitches will be scattered over several methods.
Summarizing, the backend behaves in the following way:
\begin{itemize}
\item Each flow graph is translated in a collection of methods which
  can grow dynamically. \dacom{I propose primary/secondary instead of
    the overloaded terms main/child} Each collection must contain at least one
  method, called \emph{primary}, which is the first to be created.
  All other methods, called \emph{secondary}, are added dynamically 
  whenever a new case is added to a flexswitch.

\item Each either primary or secondary method corresponds to the
  translation of some of the blocks of a single flow graph. 
\end{itemize}

When  a new case is added to a flexswitch, new blocks are generated
and translated by the backend in a new single method pointed
by a delegate which is stored in the code implementing the flexswitch,
so that the method can be invoked later.

\subsubsection{Non local links}

\commentout{

The hard bit: non-local links

@@

Moreover, the dispatch blocks are emitted only if needed, i.e. if the parent graph contains at least one flexswitch; graphs without flexswitches are rendered in the obvious way, by making one method per graph.

The slow bit: passing arguments

@@

% LocalWords:  flexswitches backend flexswitch
